% ------------------------------- Abstract ------------------------------------
\chapter*{Abstract}
\markboth{Abstract}{Abstract}

Cooperative perception lets connected vehicles extend their sensing horizon
beyond the line of sight by exchanging spatial information over
Vehicle-to-Vehicle (V2V) links. Existing benchmarks, however, almost universally
assume LiDAR as the sensing modality, leaving the feasibility of cooperation
built on low-cost \emph{passive stereo} cameras uncharacterised. This thesis
investigates whether a cooperative perception system relying exclusively on
stereo cameras can yield a measurable and characterisable improvement in
three-dimensional localisation over single-agent stereo perception, and what the
operating limits of such an approach are.

A four-stage pipeline is proposed and validated under a V-model development
process: dense stereo depth estimation, 2D object detection, lifting of
detections to 3D position, and a decentralised object-level (late) fusion layer
that registers and merges the independent observations of two cooperating
agents. Guided by the range-dependent, quadratically growing error of stereo
triangulation, the single-vehicle chain deliberately outputs 3D \emph{position}
rather than full 3D bounding boxes. A dual-dataset strategy exercises the
single-vehicle chain on real driving stereo imagery (KITTI) and the cooperative
layer on simultaneous, occlusion-truthful multi-agent recordings from the CARLA
simulator.

Empirical characterisation shows that the headline benefit of stereo
cooperation is \emph{coverage} rather than localisation: late fusion recovers
objects hidden in each agent's blind spots and raises recall substantially for
both agents symmetrically, while localisation error remains essentially flat
because the two viewpoints share a narrow viewing-angle separation that
precludes effective triangulation. The work thus establishes a baseline for
late fusion operating strictly on stereo-sourced inputs and maps the conditions
under which cooperative gains hold.

\vspace{0.6cm}
\noindent\textbf{Keywords:} cooperative perception, V2V communication, stereo
vision, depth estimation, 3D object localisation, late fusion, autonomous
driving, KITTI, CARLA.
